---
title: "Lab# - Fairness"
author: "SDS 293: Modeling for Machine Learning"
date: "Fall 2021"
output: html_document
---

# Goals for this lab
Compute trust scores for the model's predictions of wine types and handwritten numbers.
Practice working with Python in an R environment.

# Note:
The content of this lab is overwhelmingly from Google's Trust Scores paper and 
published code, just converted to work in R. The paper, code, and other resources are cited below.
This includes needing to have the trustscore.py and trustscore_evaluation.py files
in the same folder as this lab becuase the Python functions are sourced
directly from those files created by the paper's authors.

## Before we begin...

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(eval = FALSE)

```

```{r libraries}

# Load the R libraries
library(palmerpenguins) # bring in data about penguin species
library(dplyr) # tools for data wrangling in R
library(reticulate) # provides a comprehensive set of tools for interoperability between Python and R.

```

```{r}
# configure python
reticulate::py_config() # Double check that reticulate is actually using your new conda env.
reticulate::py_install("sklearn", pip = TRUE) # force install with pip. sklearn wasn't coming up via anaconda.
reticulate::py_install("matplotlib")
reticulate::py_install("keras")
reticulate::py_install("pandas")

# setting up the Python environment and bringing in the required Python packages is important.
# Running this code will likely take 5 to 10 minutes if you do not already have
# these python packages installed. Otherwise, 

# common trouble shooting:
# if you're missing a package then try adding its name in an additonal line of py_install
# if py_install isn't working then try adding the pip = TRUE argument to try installing
# the library through pip rather than anaconda
```

```{python}
import numpy as np
import trustscore # you need to have trustscore.py in the same folder as this .Rmd file to import it
import trustscore_evaluation  # you need to have trustscore_evaluation.py in the same folder as this .Rmd file to import it
import numpy as np
import matplotlib.pyplot as plt
import keras
import pandas as pd

# heads up! there might be some scary errors about "dlerror: cudart64_110.dll not found".
# It's just a warning and you can ignore it.

```

```{r}

# data
penguins_df <- penguins %>%
    dplyr::filter(complete.cases(.))
# remove all lines with missing values so the model can fit the data
 
penguins_data <- penguins_df %>%
  dplyr::select(bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g)

# target: species
penguins_target <- penguins_df %>%
  dplyr::select(species)
  
  
```


```{python}
# Import penguins from R into python
penguins_data = r.penguins_data

penguins_target = r.penguins_target

# python function: encode each penguin species to be a number.
# purpose: make it easier for the model to assign a species prediction and its trust score

dictionary = {}
current_index = 0

def encode(value):
    if value in dictionary:
      return dictionary[value]
    else:
      global current_index
      
      new_index = current_index
      dictionary[value] = new_index
      current_index = current_index + 1
      
      return new_index

X_penguins = penguins_data

y_penguins = list(map(encode, penguins_target.values.ravel()))
# at this point y_penguins is a list. We will need to make it a numpy array later.

print(dictionary) # dictionary with each penguin species connected to its assigned number
```

## Logistic Regression
```{python logistic regression model}

# NOTE: you may want to run each of these one line at a time in the console.

# import the model
from sklearn.linear_model import LogisticRegression

# Train logistic regression on digits.
model = LogisticRegression()

# fit the model
model.fit(X_penguins, y_penguins)

# Implement the model / create outputs on testing set.
y_pred = model.predict(X_penguins)
# You may get a "failed to converge" and "max number of iterations reached" warning
# this is more about the data not fitting well to the assigned model rather than a tech issue.
# it can safely be ignored.

# Initialize trust score.
trust_model = trustscore.TrustScore()

# convert the inputs into numpy arrays
X_penguins = X_penguins.to_numpy()
y_penguins = np.array(y_penguins)

# fit the trust scores model
trust_model.fit(X_penguins, y_penguins)

# the model will be fitted. No output.
```

```{python}

# Compute trusts score from the above model, given (unlabeled) testing examples and (hard) model predictions.
trust_score = trust_model.get_score(X_penguins, y_pred)

# prints the trust scores for each point in the inputted data set
print(trust_score)

```


# Check for understanding:
Why are the trust scores generally between 0 and 2?'


Are there any negative trust scores? Why do you think that is?
(Hint: think about the independent functions that create each trust score's numeric value)


Do a quick skim and find a trust score that is an outliar compared to the other numbers. Describe that trust score's numeric value through its context in the model as a whole.

# Going a step further
Try another model from the trustscore.py file, like LinearSVC or RandomForestClassifier rather LogisticRegression.
Import the model with:
from sklearn.svm import LinearSVC
from sklearn.ensemble import RandomForestClassifier


# Submitting this lab

Write about a situation in a past lab or real-life where you wish you could have used trust scores or a similar metric to check your model's predictions. If you cannot think of a situation, then detail why you don't think trust scores were necessary for that model.



```{r error tracing}
# use this code to trace back the last error, if needed.
# py_last_error()
```



# References:
Jiang, H., & Hembise, C. (n.d.). *Google/TrustScore: To trust or not to trust a classifier. A measure of uncertainty for any trained (possibly Black-box) classifier which is more effective than the classifier's own implied confidence (e.g. softmax probability for a neural network)*. GitHub. Retrieved December 22, 2021, from https://github.com/google/TrustScore

Jiang, H., Kim, B., Guan, M. Y., & Gupta, M. (2018, October 26). *To trust or not to trust a classifier*. arXiv.org. Retrieved December 22, 2021, from https://arxiv.org/abs/1805.11783 

Gajane, P., & Pechenizkiy, M. (2018, May 28). *On formalizing fairness in prediction with machine learning*. arXiv.org. Retrieved December 22, 2021, from https://arxiv.org/abs/1710.03184 

Papernot, N., & McDaniel, P. (2018, March 13). *Deep K-nearest neighbors: Towards confident, interpretable and robust deep learning*. arXiv.org. Retrieved December 22, 2021, from https://arxiv.org/abs/1803.04765v1

Interface to python. Interface to Python • reticulate. (n.d.). Retrieved December 22, 2021, from https://rstudio.github.io/reticulate/ 

Yahya, & Mortensen, P. (2021, May 8). *Convergencewarning: Lbfgs failed to converge (status=1): Stop: Total no. of iterations reached limit*. Stack Overflow. Retrieved December 22, 2021, from https://stackoverflow.com/a/62659927


# Licence:
This teachable unit is created from the references listed above, as well as my own work. It is offered under the Apache Licence.

Copyright 2021 Sarah Gillespie

Licensed under the Apache License, Version 2.0 (the "License");
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

    http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software
distributed under the License is distributed on an "AS IS" BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and
limitations under the License.

The license for the orginal trust score code, produced as part of Google research:

Copyright 2018 Google LLC

Licensed under the Apache License, Version 2.0 (the "License");
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

    https://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software
distributed under the License is distributed on an "AS IS" BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and
limitations under the License.


